Unsupervised embeddings capture latent structure of scientific mobility

### 2020-02-27 paper/current/main.tex
added text transitioning form intro into gravity law

The skip-gram embedding model preserves semantic relationships between words in a dense and meaningful low-dimensional vector-space representation~\cite{mikolov2013word2vec}, and just as its has been applied to and revealed insights implicit in text~\cite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}, so too might this model encode the latent structures of global scientific mobility. 
As scientist's careers progress, they move institutions or pick up additional affiliations, forming co-affiliation or \textit{mobility trajectories} that can be derived from their publication histories~\autocite{robinson2019mobility, sugimoto2017mostimpact}. 
We embed mobility trajectories by treating organizations analogously to words and trajectories analogously to sentences, and feed the resulting "sentence" into the standard skip-gram model (Fig.~\ref{fig:methods:mobility_sentence}).
The result is a vector-space representation of organizations in which the \textit{cosine distance} between vectors represents the tendency for pairs of organizations to occur near one another in the same career path. 
In addition to encoding latent structure, these embeddings are computationally and space efficient, robust to noise, and allow for a range of unique analyses that leverage their semantic properties using simple vector arithmetic.

We embed the trajectories of the three million name-disambiguated authors who were \textit{mobile}---having more than one affiliation---between 2008 and 2019 as evidenced by their publications indexed in the Web of Science. 
We adopt the Gravity Model of mobility~\autocite{zipf1946gravity}, one of the most common and broadly applied mobility models () to assess how well the embedding captures mobility.

>>>>

The skip-gram embedding model preserves semantic relationships between words in a dense and meaningful low-dimensional vector-space representation~\cite{mikolov2013word2vec}, and just as its has been applied to and revealed insights implicit in text~\cite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}, so too might this model encode the latent structures of global scientific mobility. 
As scientist's careers progress, they move institutions or pick up additional affiliations, forming co-affiliation or \textit{mobility trajectories} that can be derived from their publication histories~\autocite{robinson2019mobility, sugimoto2017mostimpact}. 
We embed mobility trajectories by treating organizations analogously to words and trajectories analogously to sentences, and feed the resulting "sentence" into the standard skip-gram model (Fig.~\ref{fig:methods:mobility_sentence}).
The result is a vector-space representation of organizations in which the \textit{cosine distance} between vectors represents the tendency for pairs of organizations to occur near one another in the same career path. 
In addition to encoding latent structure, these  are computationally and space efficient, robust to noise, and allow for a range of unique analyses that leverage their semantic properties using simple vector arithmetic.

We embed the trajectories of the three million name-disambiguated authors who were \textit{mobile}---having more than one affiliation---between 2008 and 2019 as evidenced by their publications indexed in the Web of Science. 
We adopt the Gravity Model of mobility~\autocite{zipf1946gravity}, one of the most common and broadly applied mobility models~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} to assess how well the embedding captures mobility. 
We derive the \textit{expected flow} from the base gravity model, defined as,

Where the flow, or number of authors co-affiliated with two organizations during the time period $f_{ij}$, increases with the organization's sizes, $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$. 
We consider two distances, $f_{ij}$, the geographic distance between the organizations, and the \textit{embedding distance} defined as the cosine distance between the organization's vector in the trained embedding space. 
We argue that the performance of a distance measure can be observed from their ability to explain expected mobility between organizations, as well as to predict flows using the gravity model. 
For geographic distance, we define $f(r_{ij})$ as the standard power-decay function, and for the embedding distance, we use the exponential form, selected as the best performing for each case.
